Scalable Extreme Deconvolution

### 2019-09-03 ml4ps_paper/paper.tex
Updated results.

We implemented both of our methods in PyTorch.
To evaluate our methods, we used a subset of rows from the Gaia DR2 source table \cite{brownGaiaDataRelease2018}.
We selected the 5 primary astrometric features, along with the BP-RP colour and mean flux in the G-band.
In total there were approximately 2 million rows.
Where data were missing, we set the field to zero and the noise variance to a large value.
This is only a small fraction of the full dataset size, but this allows us to fit the training data into memory, a requirement for use with the original implementation of extreme deconvolution.
We used the same number of components $K=128$ for each model, following \cite{andersonImprovingGaiaParallax2018}.
In practice we would want to select a value of $K$ by cross-validation.

Initialisation of all models was performed by first running a minibatch version of the K-means clustering algorithm on the dataset to get estimates of the weights and means \cite{sculleyWebscaleKmeansClustering2010}.
Initial covariances were set to the identity matrix.
Early stopping was done based on a held-out validation set comprising 10\% of the rows.
Final model performance was evaluated on a different held-out test set also comprising 10\% of the rows at the last stage, with no parameter selection performed based on this value.
Table~\ref{results-table} reports the validation and test log-likelihoods for each method.

@@

\begin{table}
  \caption{Total validation and test log-likelihoods along with average training time for the Gaia data subset. Average over 5 runs with standard deviation.}
  \label{results-table}
  \centering
  \begin{tabular}{lccc}
    \toprule
    Method     & Validation     & Test & Training Time (s) \\
    \midrule
    Existing EM & -  & -     \\
    Minibatch EM     & - & -      \\
    SGD     & $-5.468 (\pm 0.0025) \times 10^6$        & - & $1469 \pm 11$ \\
    \bottomrule
  \end{tabular}
\end{table}

>>>>

We implemented both of our methods in PyTorch.
To evaluate our methods, we used a random sample of rows from the Gaia DR2 source table~\cite{brownGaiaDataRelease2018}.
We selected the 5 primary astrometric features, along with the BP-RP colour and mean flux in the G-band.
In total there were 2 million rows.
Where data were missing, we set the field to zero and the noise variance to a large value.
This is only a small fraction of the full dataset size, but this allows us to fit the training data into memory, a requirement for use with the original implementation of extreme deconvolution.
We used the same number of components $K=128$ for each model, following \citet{andersonImprovingGaiaParallax2018}.
In practice we would want to select a value of $K$ by cross-validation.

Initialisation of all models was performed by first running a minibatch version of the K-means clustering algorithm on the dataset to get estimates of the weights and means \cite{sculleyWebscaleKmeansClustering2010}.
Initial covariances were set to the identity matrix.
Training was done for a maximum of 20 epochs.
Early stopping was done based on a held-out validation set comprising 10\% of the rows.
Final model performance was evaluated on a different held-out test set also comprising 10\% of the rows at the last stage, with no parameter selection performed based on this value.
Table~\ref{results-table} reports the validation and test log-likelihoods for each method.

@@

\begin{table}
  \caption{Total validation and test log-likelihoods along with average training time for the Gaia data subset. Average over 5 runs with standard deviation.}
  \label{results-table}
  \centering
  \begin{tabular}{lccc}
    \toprule
    Method     & Validation     & Test & Training Time (minutes) \\
    \midrule
    Existing EM & $-5.191 (\pm 0.0080) \times 10^6$  & - & $101.1 \pm 2.19$    \\
    Minibatch EM & $-5.183 (\pm 0.0005) \times 10^6$    & - & $48.7 \pm 0.39$      \\
    SGD     & $-5.193 (\pm 0.0029) \times 10^6$        & - & $36.0 \pm 0.34$ \\
    \bottomrule
  \end{tabular}
\end{table}
